---
title: "Figure 5. Relative abundances of fungal (A) and bacterial (B) taxa at the genus level "
author:"Yanmei Zhang"
data:"2024/1/1"
output: html_notebook
---

This is code to replicate the analyses and pictures from my DNA extraction method paper
Code developed by Yanmei zhang

```{r}
# How many phylum would be present after filtering? # 12
length(get_taxa_unique(ps2, taxonomic.rank="Phylum"))

# Transform into relative abundance
ps2.rel<- transform_sample_counts(ps2, function(x){x /sum(x)})

# Merge taxa at phylum level
ps2.phylum <- aggregate_taxa(ps2.rel, "Phylum")

phylum.sort <- names(sort(taxa_sums(ps2.phylum), decreasing=F))

unknown.phylum.name <- "Unknown"
phylum.name <- subset(phylum.sort, ! phylum.sort %in% unknown.phylum.name)

phylum.sort.new <- c(unknown.phylum.name, phylum.name)

# Adjust phylum label
phylum.label <- str_split_fixed(phylum.name, "__", 2)[, 2]

phylum.label.new <-c(unknown.phylum.name, phylum.label)

```

## Genus level
```{r}
# How many genera would be present after filtering? # 449 with NA included
length(get_taxa_unique(ps2,taxonomic.rank="Genus"))
 
ps2.genus.glom <- tax_glom(ps2.rel,"Genus", NArm=FALSE) ##546taxa

ps2.genus.unknown <-subset_taxa(ps2.genus.glom, is.na(Genus)) #98
filter_taxa(ps2.genus.unknown, function(x) {sum(x >= 0.05)} >= 6, TRUE) #3

ps2.genus.unknown.phy <-tax_glom(ps2.genus.unknown, "Phylum", NArm=FALSE) #7

ps2.genus.known <- subset_taxa(ps2.genus.glom, ! is.na(Genus)) #448

ps2.genus.most <- filter_taxa(ps2.genus.known, function(x) {sum(x >= 0.05)} >= 6, TRUE) #16

ps2.genus.rare <- filter_taxa(ps2.genus.known, function(x) {sum(x >= 0.05)} < 6, TRUE) #432

ps2.genus.rare.phy <- tax_glom(ps2.genus.rare, "Phylum", NArm=FALSE) #11

ps2.mix <- merge_phyloseq(ps2.genus.unknown.phy, ps2.genus.rare.phy, ps2.genus.most) #34

tax4 <-  as.data.frame(tax_table(ps2.mix))

tax4 <- tax4 %>%
  dplyr::mutate(mix = tax4$Genus, .before = "Phylum")

tax4 <- tax4 %>%
  mutate(mix = replace(mix, is.na(mix), Phylum[is.na(mix)])) 

tax4 <- as.matrix(tax4)
ps2.mix1 <- merge_phyloseq(ps2.mix, tax_table(tax4)) #34

ps2.mix.glom <- aggregate_taxa(ps2.mix1, "mix") #28=16+12
```


Make adjustment of order of genus name 
```{r}
genus.sort <- names(sort(taxa_sums(ps2.mix.glom), decreasing=F))

unknown.phylum.name <- "Unknown"
phylum.name <-  str_subset(genus.sort, pattern ="p__")
phylum.name <- phylum.sort.new 

genus.name <-  str_subset(genus.sort , pattern = "g__") 

genus.name.A <- subset(genus.name, genus.name %in% 
                         tax_table(ps2.mix.glom)[tax_table(ps2.mix.glom)[, "Phylum"] == 
                                                   "p__Ascomycota", "mix"])
genus.name.B <- subset(genus.name, genus.name %in% 
                         tax_table(ps2.mix.glom)[tax_table(ps2.mix.glom)[, "Phylum"] == 
                                                   "p__Basidiomycota", "mix"])
genus.sort.new <- c(phylum.name,genus.name.A, genus.name.B)

# Adjust genus label
phylum.label <- phylum.label.new
phylum.label[c(11, 12)] <- paste("other", phylum.label[c(11, 12)], sep = " ")

genus.label.A <- paste(genus.name.A, "(A)", sep = " ")
genus.label.B <- paste(genus.name.B, "(B)", sep = " ")

genus.label.A <- str_split_fixed(genus.label.A, "__", 2)[, 2]
genus.label.B  <- str_split_fixed(genus.label.B , "__", 2)[, 2]

genus.label.new <- c(phylum.label, genus.label.A, genus.label.B)

# Table for ggplot
genus.mix <- otu_table(ps2.mix.glom) %>% data.frame(check.names = F)
genus.mix$mix <- factor(rownames(genus.mix), levels = genus.sort.new, labels = genus.label.new)
genus.mix <- melt(genus.mix, id ="mix")
names(genus.mix)[2]<- "SampleID"
names(genus.mix)[3]<- "Relative_abundance"

# Map info
group <- sample_data(ps2.mix.glom) %>% data.frame(check.names = F)
group$SampleID <- row.names(group)
genus.mix_samp<- merge(genus.mix, group, by = 'SampleID')

genus.mix_samp$Protocol <- factor(genus.mix_samp$Protocol, level = c("QIAGEN", "CTAB"))
```

Composition barplots-fungi
```{r}
n <- 28

# Make a color vector
palette_6 = pal_d3("category20c")(6)
show_col(palette_6)

palette_10= pal_d3("category10")(10)
show_col(palette_10)

palette_28 <- c(palette_12_1, palette_6, palette_10)
show_col(palette_28)

p.genus <- ggplot(genus.mix_samp, aes(as.factor(as.character(Year_since_decay_start)),
                                100*Relative_abundance, fill = mix)) +
  geom_col(position = 'stack', width = 0.85) +
  facet_grid(Tree_log ~ Protocol, scales = 'free_x') +
  scale_fill_manual(values =  palette_28) +
  labs(x = 'Year_since_decay_start', y = 'Relative Abundance (%)') +
  guides(fill = guide_legend(ncol = 2)) +
  theme_minimal(9) +
  theme(axis.title.x = element_blank(),
        legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
p.genus

source("src/data_summary.R")
genus.mix_samp_sum <- data_summary(genus.mix_samp, varname="Relative_abundance", 
                              groupnames=c("Tree_host","Protocol", 
                                           "Year_since_decay_start", "mix"))


p.genus <- ggplot(genus.mix_samp_sum, aes(as.factor(as.character(Year_since_decay_start)),
                                    100*Relative_abundance, fill = mix)) +
  geom_col(position = 'stack', width = 0.85) +
  facet_grid(Tree_host~Protocol, scales = 'free_x') +
  scale_fill_manual(values =  palette_28) +
  labs(x = 'Year_since_decay_start', y = 'Relative Abundance (%)') +
  guides(fill = guide_legend(ncol = 2)) +
  theme_minimal(9) +
  theme(axis.title.x = element_blank(),
        legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
p.genus
```


## For bacteria, too much phylum in the picture so we pick up top 11 phylum and assign the rest as rare phylum
```{r}
phylum.sort <- names(sort(taxa_sums(ps2.phylum), decreasing=F))
phylum.top11 <- phylum.sort[23:33]

phylum.rare <- subset(phylum.sort, ! phylum.sort %in% phylum.top11)

ps2.phylum.new <- ps2.phylum
tax_table(ps2.phylum.new)[tax_table(ps2.phylum.new)[,"Phylum"] %in% phylum.rare, 
                         "Phylum"] <- "rare phylum"

# Remove Kingdom
tax_table(ps2.phylum.new) = tax_table(ps2.phylum.new)[, -1]
ps2.phylum.new1 <- aggregate_taxa(ps2.phylum.new, "Phylum") #12


rare.phylum.name <- "rare phylum"

phylum.sort.new <- c(rare.phylum.name, phylum.top11)
```

```{r}
ps2.mix.glom.new <- ps2.mix.glom
tax_table(ps2.mix.glom.new)[tax_table(ps2.mix.glom.new)[,"mix"] %in% phylum.rare, 
                            "mix"] <- "rare phylum" 
tax_table(ps2.mix.glom.new) = tax_table(ps2.mix.glom.new)[, -1]

ps2.mix.glom.new1 <- aggregate_taxa(ps2.mix.glom.new, "mix")
```

Make adjustment of order of genus name 
```{r}
genus.sort <- names(sort(taxa_sums(ps2.mix.glom.new1), decreasing=F))

rare.phylum.name <- "rare phylum"

phylum.name <-  subset(genus.sort, genus.sort %in% phylum.top11)


genus.name <- subset(genus.sort, ! genus.sort %in% c(rare.phylum.name,phylum.name))

genus.sort.new <- c(rare.phylum.name,phylum.name, genus.name)

# Table for ggplot
genus.mix <- otu_table(ps2.mix.glom.new1) %>% data.frame(check.names = F)
genus.mix$mix <- factor(rownames(genus.mix), levels = genus.sort.new)
genus.mix <- melt(genus.mix, id ="mix")
names(genus.mix)[2]<- "SampleID"
names(genus.mix)[3]<- "Relative_abundance"

# Map info
group <- sample_data(ps2.mix.glom.new1) %>% data.frame(check.names = F)
group$SampleID <- row.names(group)
genus.mix_samp<- merge(genus.mix, group, by = 'SampleID')

genus.mix_samp$Protocol <- factor(genus.mix_samp$Protocol, level = c("QIAGEN", "CTAB"))
```

Composition barplots-bacteria
```{r}
# Make a color vector
n <- 37

palette_15 = pal_d3("category20c")(15)
show_col(palette_15)

palette_10 = pal_d3("category10")(10)
show_col(palette_10)

palette_37 <- c(palette_12_1, palette_15, palette_10)
show_col(palette_37)

p.genus <- ggplot(genus.mix_samp, aes(as.factor(as.character(Year_since_decay_start)),
                                100*Relative_abundance, fill = mix)) +
  geom_col(position = 'stack', width = 0.85) +
  facet_grid(Tree_log ~ Protocol, scales = 'free_x') +
  scale_fill_manual(values =  palette_37) +
  labs(x = 'Year_since_decay_start', y = 'Relative Abundance (%)') +
  guides(fill = guide_legend(ncol = 2)) +
  theme_minimal(9) +
  theme(axis.title.x = element_blank(),
        legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
p.genus

genus.mix_samp_sum <- data_summary(genus.mix_samp, varname="Relative_abundance", 
                              groupnames=c("Tree_host","Protocol", 
                                           "Year_since_decay_start", "mix"))


p.genus <- ggplot(genus.mix_samp_sum, aes(as.factor(as.character(Year_since_decay_start)),
                                    100*Relative_abundance, fill = mix)) +
  geom_col(position = 'stack', width = 0.85) +
  facet_grid(Tree_host~Protocol, scales = 'free_x') +
  scale_fill_manual(values =  palette_37) +
  labs(x = 'Year_since_decay_start', y = 'Relative Abundance (%)') +
  guides(fill = guide_legend(ncol = 2)) +
  theme_minimal(9) +
  theme(axis.title.x = element_blank(),
        legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
p.genus
```
