---
title: "Figure 3. Read count summary for amplicon sequencing data"
author:"Yanmei Zhang"
data:"2024/1/1"
output: html_notebook
---

### Track sample reads depth
Let's see the tracking summary of sequencing depth during reads processing
```{r}
# Loading reads track file
## for fungi
dada_track = read.delim("../03_tabletax/dada2_its2_track.txt")
## for bacteria, change it into
dada_track = read.delim("../02_tabletax/dada2_16s_track.txt")

dada_track <-  dada_track %>% mutate(SampleID = row.names(.))

depth_track5 <- merge(dada_track, 
                      depth_track4[, c("SampleID", "Tree_log", "Tree_host", "Protocol",
                                       "Year_since_decay_start", "DNA_yields", "nonbacterial_filtered",
                                       "decontam", "quality_filtered")],
              by = "SampleID",all = F)


track_l<- reshape2::melt(depth_track5, id.vars=c("SampleID", "Tree_log", "Tree_host", "Protocol",
                                                 "Year_since_decay_start", "DNA_yields"),
                         variable.name = "step", value.name = "reads_count")

track_l$Protocol<-factor(track_l$Protocol, level = c("QIAGEN", "CTAB"))
head(track_l)

p <-  ggplot(track_l,aes(x = step, y = reads_count, color = Protocol)) +
  geom_line(aes(group = SampleID, color = Protocol)) +
  geom_point(aes(color= Protocol), pch = 21,size = 1) +
  facet_grid(Year_since_decay_start~Tree_host) +
  scale_color_manual(values =  c("#F8766D", "#00BA38")) +
  scale_fill_manual(values =  c("#F8766D", "#00BA38"))

p <-  p +
  theme_bw(7)+
  theme(legend.key.size = unit(0.2, "cm"),
                    legend.title = element_blank()) +
  theme(axis.title.x = element_blank())


if (length(unique(track_l$SampleID))>3){  
  p=p+theme(axis.text.x=element_text(angle=45,vjust=1, hjust=1))
}

p
```

### t test to test the effect of DECAY STATUS on reading depth adter filtration
The goal is to assess the effect of decay status on the read depth. The samples are not paired
```{r}
# Decay status effect
depth_track4 <- depth_track3 %>%
  filter(Sample_or_Control == "Sample") %>%
  mutate(status = ifelse(Year_since_decay_start %in% c(0), "non_decayed", "decayed")) 

depth_track4$status <- factor(depth_track4$status, level = c("non_decayed", "decayed"))
depth_track4 %>%
  group_by(status) %>%
  get_summary_stats(quality_filtered, show = c("mean","sd", "min", "max"))

# Check assumptions
## Outliers
depth_track4 %>%
  group_by(status) %>%
  identify_outliers(quality_filtered)
# no outliers were found
## Normality
depth_track4 %>%
  group_by(status) %>%
  shapiro_test(quality_filtered)
# Non-decayed group not normally distributed
ggqqplot(depth_track4, "quality_filtered", facet.by = "status")
# Or build the linear model
model  <- lm(quality_filtered ~ status, data = depth_track4)
# Create a QQ plot of residuals
ggqqplot(residuals(model))
# Compute Shapiro-Wilk test of normality
shapiro_test(residuals(model))
# The data is not normally distributed

# Computation
stat.test <- depth_track4 %>% 
  wilcox_test(quality_filtered ~ status) %>%
  add_significance()
stat.test

# Effect size
depth_track4 %>% wilcox_effsize(quality_filtered ~ status)

stat.test <- stat.test %>% add_xy_position(x = "status")

ggplot(depth_track4, aes(x=status, y=quality_filtered, color=status))+
  geom_boxplot(aes(fill = status), alpha = 0.3)+
  geom_jitter(position = position_jitterdodge(),size =1, alpha = 0.3) +
  #scale_y_log10()+
  #facet_grid(Tree_host~.)+
  scale_color_manual(values=c("lightgrey", "#5C4033")) +
  scale_fill_manual(values=c("lightgrey", "#5C4033")) +
  labs(x = "Decay Status", y = "Quality Filtered Reads")+
  theme_bw(7) +
  theme(legend.key.size = unit(0.2, "cm"),
                    legend.title = element_blank()) +
  stat_pvalue_manual(stat.test, tip.length = 0, hide.ns = TRUE)
