---
title: "Figure 1. DNA extraction yields from sound or decayed wood"
author:"Yanmei Zhang"
data:"2024/1/1"
output: html_notebook
---

This is code to replicate the analyses and pictures from my DNA extraction method paper
Code developed by Yanmei zhang

## Data pre-processing
```{r}
input_file <- "DNA_quantification.csv"
data=read.csv(input_file, header=T)
str(data)

data <- data %>%
  convert_as_factor(Tree_log, Tree_host, Year_since_decay_start, Protocol, When, Year)
data$Protocol=factor(data$Protocol, levels = c('QIAGEN', 'CTAB', 'SDS')) 
data$Year_since_decay_start<- factor(data$Year_since_decay_start, 
                       levels = c(0, 1, 7),
                       labels = c("0", "1", "7"))
str(data)
```

## Friedman Test
As the data is not normally distributed, I will try use the Friedman test. The Friedman test is a non-parametric alternative to the one-way repeated measures ANOVA test. It extends the Sign test in the situation where there are more than two groups to compare.
```{r}
# Visualization
bxp <- ggboxplot(
  data, x = "Protocol", y = "DNA_yields", color = "Protocol", fill ="Protocol",
  add = "jitter", add.params = list(size=1, alpha=0.3), alpha =0.3,
  short.panel.labs = FALSE) +
  labs(y="DNA_yields (ng/mg)"
       )  

res.fried <- data %>% friedman_test(DNA_yields ~ Protocol |Sample_name)

data%>% friedman_effsize(DNA_yields ~ Protocol |Sample_name)

pwc <- data %>%
  wilcox_test(DNA_yields ~ Protocol , paired = TRUE, p.adjust.method = "bonferroni")
pwc

# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "Protocol")

bxp3 <- bxp + 
  stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
  labs(subtitle = get_test_label(res.fried, detailed = TRUE), caption = get_pwc_label(pwc)) +
  theme_bw(7) +
  theme(legend.key.size = unit(0.2, "cm"),
                    legend.title = element_blank())
```

## Two way repeated anova
A two-way repeated measures ANOVA was performed to evaluate the effect of different extraction protocols over decay time on DNA yield for birch and pine, respectively.
### Birch
```{r}
###----------------------------Subset birch samples-------------------------------------###
# Subset birch samples
Onlybirch<-subset(data.long, Tree_host=="paper_birch")

###-----------------------------Summary statistics-------------------------------------###
Onlybirch %>%
  group_by(Year_since_decay_start, Protocol) %>%
  get_summary_stats(DNA_yields, type = "mean_sd")

###-----------------------------Data visualization-------------------------------------###
# Create box plots of DNA yields by year since decay start, colored by protocols:
bxp.birch <- ggboxplot(
  Onlybirch, x = "Year_since_decay_start", y = "DNA_yields", color = "Protocol", 
  add = "jitter", add.params = list(size=1, alpha=0.3)) + 
  labs(y="DNA_yields (ng/mg)") 

###-----------------------------Two-way repeated ANOVA-------------------------------###
# 2 within- subjects variables
# Check assumptions
# 1.Outliers
Onlybirch %>%
  group_by(Year_since_decay_start, Protocol) %>%
  identify_outliers(DNA_yields)
## there were 3 extreme outliers (C25, C29, S25).

# 2.Normality assumption
Onlybirch %>%
  group_by(Year_since_decay_start, Protocol) %>%
  shapiro_test(DNA_yields) 
## The DNA yield was normally distributed (p > 0.05), except for 2 cells
## as assessed by Shapiro-Wilk’s test of normality.

ggqqplot(Onlybirch, "DNA_yields", ggtheme = theme_bw()) +
  facet_grid(Protocol ~ Year_since_decay_start, labeller = "label_both")

# 3. Assumption of sphericity
# Computation
res.aov.b <- anova_test(
  data = Onlybirch, dv = DNA_yields, wid = Tree_log,
  within = c(Year_since_decay_start, Protocol)
)
get_anova_table(res.aov.b)

# Post-hoc tests
# Procedure for a significant two-way interaction
# 1. Simple main effect
# Effect of protocols at each time point
one.way.b <- Onlybirch %>%
  group_by(Year_since_decay_start) %>%
  anova_test(dv = DNA_yields, wid = Tree_log, within = Protocol) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way.b

# Pairwise comparisons between protocol groups
pwc.b <- Onlybirch %>%
  group_by(Year_since_decay_start) %>%
  pairwise_t_test(
    DNA_yields ~ Protocol, paired = TRUE,
    p.adjust.method = "bonferroni"
  )
pwc.b

# Visualization: box plots with p-values
pwc.b <- pwc.b %>% add_xy_position(x = "Year_since_decay_start")
bxp.birch1 <- bxp.birch + 
  stat_pvalue_manual(pwc.b, tip.length = 0, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.aov.b, detailed = TRUE),
    caption = get_pwc_label(pwc.b)
  )+
  theme_bw(7) +
  theme(legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
bxp.birch1
```

### Pine 
```{r}
###----------------------------Subset pine samples-------------------------------------###
# Subset pine samples
Onlypine<-subset(data.long, Tree_host=="red_pine")

###-----------------------------Summary statistics-------------------------------------###
Onlypine %>%
  group_by(Year_since_decay_start, Protocol) %>%
  get_summary_stats(DNA_yields, type = "mean_sd")

###-----------------------------Data visualization-------------------------------------###
# Create box plots of DNA yields by year since decay start, colored by protocols:
bxp.pine <- ggboxplot(
  Onlypine, x = "Year_since_decay_start", y = "DNA_yields", color = "Protocol", 
  add = "jitter", add.params = list(size=1, alpha=0.3)) +
  labs(y="DNA_yields (ng/mg)") 
bxp.pine

###-----------------------------Two-way repeated ANOVA-------------------------------###
# 2 within- subjects variables
# Check assumptions
# 1.Outliers
Onlypine %>%
  group_by(Year_since_decay_start, Protocol) %>%
  identify_outliers(DNA_yields) 
## there were 2 extreme outliers (S52, S57).

# 2.Normality assumption
Onlypine%>%
  group_by(Year_since_decay_start, Protocol) %>%
  shapiro_test(DNA_yields) 
## The DNA yield was normally distributed (p > 0.05), except for 3 cells
## as assessed by Shapiro-Wilk’s test of normality.

ggqqplot(Onlypine, "DNA_yields", ggtheme = theme_bw()) +
  facet_grid(Protocol ~ Year_since_decay_start, labeller = "label_both")

# 3. Assumption of sphericity
# Computation
res.aov.p <- anova_test(
  data = Onlypine, dv = DNA_yields, wid = Tree_log,
  within = c(Year_since_decay_start, Protocol)
)
get_anova_table(res.aov.p)

# Post-hoc tests
# Procedure for a significant two-way interaction
# 1. Simple main effect
# Effect of protocols at each time point
one.way.p <- Onlypine %>%
  group_by(Year_since_decay_start) %>%
  anova_test(dv = DNA_yields, wid = Tree_log, within = Protocol) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way.p

# Pairwise comparisons between protocol groups
pwc.p <- Onlypine %>%
  group_by(Year_since_decay_start) %>%
  pairwise_t_test(
    DNA_yields ~ Protocol, paired = TRUE,
    p.adjust.method = "bonferroni"
  )
pwc.p

# Visualization: box plots with p-values
pwc.p <- pwc.p %>% add_xy_position(x = "Year_since_decay_start")
bxp.pine1 <- bxp.pine + 
  stat_pvalue_manual(pwc.p, tip.length = 0, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.aov.p, detailed = TRUE),
    caption = get_pwc_label(pwc.p)
  ) +
  theme_bw(7) +
  theme(legend.key.size = unit(0.2, "cm"),
        legend.title = element_blank()) 
bxp.pine1
```